---
title: "R Notebook"
out: html_notebook
---

```{r Load libraries, message=FALSE, warning=FALSE, include=FALSE}
library(readr)
library(tidyverse)
library(skimr)
library(lubridate)
library(janitor)
library(ggrepel)
library(zoo)
```

```{r load data, include=FALSE}
load("~/Desktop/prospectus/data/ripa/person.RData")
load("~/Desktop/prospectus/data/ripa/incident.RData")
```


```{r}
ripa = merge(person, incident)
```


#Drop Native American + other
```{r}
ripa$race = as.character(ripa$race)
```

```{r}
ripa <- filter(ripa, !race == "Native")
```



#Creates male/female variables
```{r}
ripa$male = 0
ripa$male[ripa$gender == "1"] <- 1
ripa$female = 0
ripa$female[ripa$gender == "0"] <- 1
```



```{r}
ripa$gender[ripa$gender == "0"] <- "Female"
ripa$gender[ripa$gender == "1"] <- "Male"

ripa$gender = as.factor(ripa$gender)
```


# Changes so3 from 0,1 to Before/After
```{r}
ripa$so3[ripa$so3 == 0] <- "Before"
ripa$so3[ripa$so3 == 1] <- "After"
```

```{r}
ripa$so3 = as.factor(ripa$so3)
```

#Drops variables 
```{r}
ripa$action01_order <- NULL
ripa$action02_contact1 <- NULL
ripa$action03_test <- NULL
ripa$action04_curbside <- NULL
ripa$action05_cuffed <- NULL
ripa$action06_detention <- NULL
ripa$action07_canine1 <- NULL
ripa$action08_firearm1 <- NULL
ripa$action09_firearm2 <- NULL
ripa$action11_device <- NULL
ripa$action12_projectile <- NULL
ripa$action13_canine2 <- NULL
ripa$action14_baton <- NULL
ripa$action15_chemical <- NULL
ripa$action16_contact1 <- NULL
ripa$action17_photo <- NULL
ripa$action18_ask_search_pe <- NULL
#ripa$action19_search_pe
#ripa$action20_search_pr
ripa$action21_seized <- NULL
ripa$action22_impound <- NULL
#ripa$action_taken
ripa$action22_none <- NULL

ripa$a18_consent_granted <- NULL
ripa$r1_noaction <- NULL
ripa$r2_warning <- NULL
ripa$r3_citation <- NULL
ripa$r4_cite_release <- NULL
ripa$r5_arrest1 <- NULL
ripa$r6_arrest2 <- NULL
ripa$r7_interview <- NULL
ripa$r8_transport <- NULL
ripa$r9_guardian <- NULL
ripa$r10_psych <- NULL
ripa$r11_refUSDHS <- NULL
ripa$r12_refschool1 <- NULL
ripa$r13_refschool2 <- NULL
ripa$action20_ask_search_pr <- NULL
ripa$a20_consent_granted <-NULL
ripa$action19_search_pe <- NULL
ripa$action20_search_pr <- NULL
ripa$c1_firearms <- NULL
ripa$c2_ammunition <- NULL
ripa$c3_weapons <- NULL
ripa$c4_drugs <- NULL
ripa$c5_alcohol <- NULL
ripa$c6_money <- NULL
ripa$c7_paraphernalia <- NULL
ripa$c8_stolen <- NULL
ripa$c9_devices <- NULL
ripa$c10_other <- NULL
ripa$k12 <- NULL
ripa$school <- NULL
```


#Drops observations with missing data on gender

```{r}
ripa$gender = as.character(ripa$gender)
```

```{r}
ripa <- filter(ripa, !is.na(ripa$gender))
```

#Drops variable gender

```{r}
ripa$gender <- NULL
```




#Creates vector with variables names 
```{r}
ripa_vars <- c("person_id", "form_id", "persons", "so3","duration",  
               "year", "yearmonth", "date",  "stop_time", 
               "bureau", "n_division", "division", "rd",  "cfs",
               "officer1", "officer2", "officer_div", "officer_asgmt",
               "stop_type", "offense_code", "offense_type", 
               "rfs","trf", "rs", "other_stops", "trf_violation", "brs", "race",
               "age", "male", "female",  
               "asian", "black", "hispanic", "white", 
               "search_person", "search_property", "hit", "contraband_none", 
               "ros", "ros1_arrest", "ros2_citation", "ros3_interview", "ros4_naw", "ros5_other",
               "action_taken")       
```

                        
 
# Subset variables
```{r}
ripa <- ripa[, ripa_vars]
```



# Creates variable week
```{r}
ripa$week = week(ripa$date)
```



# Orders levels of variables race
```{r}
ripa$race = factor(ripa$race, levels = c("Asian", "Black", "Hispanic", "White"))
```


# Drops observations with reason for stop == "Other"
```{r}
ripa$rfs = as.character(ripa$rfs)
```

```{r}
ripa <- filter(ripa, !(rfs == "Other"))
```



# Explores missing data in variable stop_type (Vehicular, Pedestrian and Cyclist)
```{r}
ripa$type = "Reported 6_30_2021"
ripa$type[ripa$stop_type == "driver"] <- "vehicular"
ripa$type[ripa$stop_type == "passenger"] <- "vehicular"
ripa$type[ripa$stop_type == "pedestrian"] <- "pedestrian"
ripa$type[ripa$stop_type == "bicycle"] <- "cyclist"
```




# Drops Cyclist and Pedestrian stops
```{r}
ripa <- filter(ripa, !type == "cyclist")
ripa <- filter(ripa, !type == "pedestrian")
```



# Changes rfs "Reasonable suspicion" to "Suspicion"
```{r}
ripa$rfs[ripa$rfs == "Reasonable suspicion"] <- "Suspicion"
```



#Changes levels of variables rfs 
```{r}
ripa$rfs = factor(ripa$rfs, levels = c("Traffic", "Suspicion"))
```



# Imputes type of stop as vehicular when the variable "traffic violation" is populated. This includes all moving, non moving and equipment violations
```{r}
ripa$type[ripa$trf_violation == "moving"] <- "vehicular"
ripa$type[ripa$trf_violation == "non-moving"] <- "vehicular"
ripa$type[ripa$trf_violation == "equipment"] <- "vehicular"
```


# Drops observatiopns of unidentified type reported before 6_30_2021
```{r}
ripa <- filter(ripa, !type == "Reported 6_30_2021")

ripa
```

# Orders levels of variable so3
```{r}
ripa$so3 = factor(ripa$so3, levels = c("Before", "After"))
```


# Creates dummy variables for each race
```{r}
ripa$race = NA
ripa$race[ripa$asian == 1] <- "Asian"
ripa$race[ripa$black == 1] <- "Black"
ripa$race[ripa$hispanic == 1] <- "Hispanic"
ripa$race[ripa$white == 1] <- "White"
```



```{r}
ripa$age_group = NA
ripa$age_group[ripa$age < 18] <- "Up to 17"
ripa$age_group[ripa$age == 18] <- "18 to 25"
ripa$age_group[ripa$age > 18 & ripa$age < 26 ] <- "18 to 25"
ripa$age_group[ripa$age > 25 & ripa$age < 31 ] <- "26 to 30"
ripa$age_group[ripa$age > 30 & ripa$age < 36 ] <- "31 to 35"
ripa$age_group[ripa$age > 35  & ripa$age < 41] <- "36 to 40"
ripa$age_group[ripa$age > 40 & ripa$age < 46] <- "41 to 45"
ripa$age_group[ripa$age > 45 & ripa$age < 51] <- "46 to 50"
ripa$age_group[ripa$age > 50  & ripa$age < 56 ] <- "51 to 55"
ripa$age_group[ripa$age > 55 & ripa$age < 61 ] <- "56 to 60"
ripa$age_group[ripa$age > 60 & ripa$age < 66 ] <- "61 to 65"
ripa$age_group[ripa$age > 65 & ripa$age < 71] <- "66 to 70"
ripa$age_group[ripa$age > 70 & ripa$age < 76] <- "71 to 75"
ripa$age_group[ripa$age > 75  & ripa$age < 81 ] <- "76 to 80"
ripa$age_group[ripa$age > 80] <- "More than 80"

```




```{r}
ripa$age_group = factor(ripa$age_group, levels = c("Up to 17", 
                                                   "18 to 25",
                                                   "26 to 30",
                                                   "31 to 35",
                                                   "36 to 40",
                                                   "41 to 45",
                                                   "46 to 50",
                                                   "51 to 55",
                                                   "56 to 60",
                                                   "61 to 65",
                                                   "66 to 70",
                                                   "71 to 75",
                                                   "76 to 80",
                                                   "More than 80"
                                                   ))
```


```{r}
ripa$gender = NA

ripa$gender[ripa$male == 1] <- "Male"
ripa$gender[ripa$male == 0] <- "Female"

ripa$gender = factor(ripa$gender, levels = c("Male", "Female"))
```


```{r}
ripa$race = factor(ripa$race, levels = c("Asian", "Black", "Hispanic", "White"))
```


```{r}
ripa$hit_person = 0
ripa$hit_person[ripa$search_person == 1 & ripa$hit == 1] <- 1

```

```{r}
ripa$hit_property = 0
ripa$hit_property[ripa$search_property == 1 & ripa$hit == 1] <- 1
```


```{r}
table(ripa$ros, useNA = "ifany")
```


```{r}
save(ripa, file = "~/Desktop/prospectus/data/ripa/ripa.RData")

write.csv(ripa, file = "~/Desktop/prospectus/data/ripa/ripa.csv")
```



```{r}
table(ripa$rfs, useNA = "ifany")
```





